Composable Concurrency Models

### 2016-11-18 slides/slides.tex
Typeset atomic variable figures
Typeset remaining figures and started fixing conclusion

\usepackage{appendixnumberbeamer}
\usepackage[USenglish]{babel}
\usepackage[backend=bibtex,citestyle=authortitle]{biblatex}
\usepackage[scale=2]{ccicons}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{multicol}
\usepackage{tikz}

@@

@@

    \begin{columns}
      \begin{column}{0.5\textwidth}
        \centering
        \includegraphics[width=160pt]{../figures/nonatomic}
      \end{column}
      \hfill
      \begin{column}{0.5\textwidth}
        $x = 23$

@@

    \begin{columns}
      \begin{column}{0.5\textwidth}
        \centering
        \includegraphics[width=160pt]{../figures/atomic}
      \end{column}
      \hfill
      \begin{column}{0.5\textwidth}
        $x = 23$

@@

  \begin{frame}
    \frametitle{Atomic variables}
    \centering

    \includegraphics[width=220pt]{../figures/lock}
  \end{frame}

  \begin{frame}
    \frametitle{Atomic variables}
    \centering

    \includegraphics[width=220pt]{../figures/multiple-locks}
  \end{frame}

@@

  \begin{frame}
    \frametitle{Software transactional memory}
    \centering

    \includegraphics[width=220pt]{../figures/stm}
  \end{frame}

@@

  \begin{frame}
    \frametitle{Communicating threads}
    \centering

    \includegraphics[width=220pt]{../figures/agent}
    \vfill
    \textbf{Agents:} An isolated thread handles all operations on an object.~\footcite{Swalens2014}
  \end{frame}

  \begin{frame}
    \frametitle{Communicating threads}
    \centering

    \includegraphics[width=220pt]{../figures/csp}
    \vfill
    \textbf{Communicating sequential processes (CSP):} Independent threads communicate synchronously through predefined channels.~\footcite{Hoare1978}
  \end{frame}

  \begin{frame}
    \frametitle{Communicating threads}
    \centering

    \includegraphics[width=220pt]{../figures/actor}
    \vfill
    \textbf{The actor model:} Independent threads send messages to known addresses.~\footcite{Agha1986}
  \end{frame}

@@

  \begin{frame}
    \frametitle{Composability criteria}
    \centering

    \includegraphics[width=360pt]{../figures/clojure-survey}
  \end{frame}

@@

  \subsection{Conclusions}

  \begin{frame}
    \frametitle{Conclusions}

    \begin{itemize}
      \onslide<1->{
        \item The language may ensure that some bad interactions don't happen.
      }
      \vfill
      \onslide<2->{
        \item Sometimes, it's best to just use a different concurrency model.
      }
      \vfill
      \onslide<3->{
        \item Future work would identify common ``building blocks'' (thread creation, message passing, etc.) that could be composed to express existing concurrency models and combinations of models.
      }
    \end{itemize}
  \end{frame}

>>>>

\usepackage{appendixnumberbeamer}
\usepackage[USenglish]{babel}
\usepackage[backend=bibtex,citestyle=authortitle]{biblatex}
\usepackage[scale=2]{ccicons}
\usepackage{color}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{multicol}
\usepackage{pifont}
\usepackage{tikz}
\usetikzlibrary{positioning}

@@

  \section*{Introduction}

@@

    \begin{columns}
      \begin{column}{0.5\textwidth}
        \centering
        \input{../figures/nonatomic.tikz}
      \end{column}
      \hfill
      \begin{column}{0.5\textwidth}
        $x = 23$

@@

    \begin{columns}
      \begin{column}{0.5\textwidth}
        \centering
        \input{../figures/atomic.tikz}
      \end{column}
      \hfill
      \begin{column}{0.5\textwidth}
        $x = 23$

@@

  \begin{frame}
    \frametitle{Atomic variables}
    \centering
    \input{../figures/lock.tikz}
  \end{frame}

  \begin{frame}
    \frametitle{Atomic variables}
    \centering
    \input{../figures/multiple-locks.tikz}
  \end{frame}

@@

  \begin{frame}
    \frametitle{Software transactional memory}
    \centering
    \input{../figures/stm.tikz}
  \end{frame}

@@

  \begin{frame}
    \frametitle{Communicating threads}
    \centering
    \input{../figures/agent.tikz}
    \vfill
    \textbf{Agents:} An isolated thread wraps an object.~\footcite{Swalens2014}
  \end{frame}

  \begin{frame}
    \frametitle{Communicating threads}
    \centering
    \input{../figures/csp.tikz}
    \vfill
    \textbf{Communicating sequential processes (CSP):} Independent threads communicate synchronously through predefined channels.~\footcite{Hoare1978}
  \end{frame}

  \begin{frame}
    \frametitle{Communicating threads}
    \centering
    \input{../figures/actor.tikz}
    \vfill
    \textbf{The actor model:} Independent threads send messages to known addresses.~\footcite{Agha1986}
  \end{frame}

@@

  \begin{frame}
    \frametitle{Composability criteria}
    \centering
    \small

    \newcommand{\n}{\textcolor{red}{\ding{55}}}
    \newcommand{\y}{\textcolor{olive}{\ding{51}}}

    \textbf{Pairwise combinations of four Clojure concurrency models:}
    \vfill
    \begin{tabular}{c | c c c c | c c c c}
      & \multicolumn{4}{c|}{\textbf{Safety}} & \multicolumn{4}{|c}{\textbf{Liveness}} \\
      & atoms & refs & agents & channels & atoms & refs & agents & channels \\
      \hline
      atoms & \n & \n & \n & \n & \y & \y & \y & \n \\
      refs & \n & \y & \y & \n & \y & \y & \y & \n \\
      agents & \y & \y & \y & \y & \y & \y & \y & \n \\
      channels & \y & \y & \y & \y & \y & \y & \n & \n
    \end{tabular}
  \end{frame}

@@

  \subsection{Future work}

  \begin{frame}
    \frametitle{Future work}

    \begin{itemize}
      \onslide<1->{
        \item Swalens et al.\ planned to identify composable ``building blocks'' (thread creation, message passing, etc.) that could be used to build common concurrency models.~\footcite{Swalens2014}
      }
      \vfill
      \onslide<2->{
        \item Marr and D'Hondt proposed a unifying abstraction for high-level language virtual machines.~\footcite{Marr2012}
      }
      \vfill
      \onslide<3->{
        \item Ziv et al.\ formalized a theory for safely composing concurrency control.~\footcite{Ziv2015}
      }
    \end{itemize}
  \end{frame}
